# Quick start {#quick-start}

```{r include=FALSE}
library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE, eval = TRUE, echo = TRUE, cache = TRUE)
library(genekitr)
library(dplyr)
library(DT)
```

## [Search gene metadata](#search-gene-metadata-1)

```{r eval=FALSE}
id <- c("MCM10", "CDC20", "S100A9", "MMP1", "BCC7","FAKEID","HBD", "NUDT10")
info <- genInfo(id, org = "hs")
# keep unique records
info2 <- genInfo(id, org = "hs", unique = TRUE)
```


## [Search PubMed records](#search-pubmed-1)

```{r eval=FALSE}
## check package version
packageVersion("genekitr")

## for version < 0.8.6
term <- c("Tp53", "Brca1", "Tet2")
keys <- c("stem cell", "mouse")
l <- getPubmed(term, keys)

## for version > 0.8.6
term <- c("Tp53", "Brca1", "Tet2")
add_term <- c("stem cell", "mouse")
l <- getPubmed(term, add_term, num = 30)

# set "num" larger to get all records
all <- getPubmed(term, add_term, num = 666666)
```

## [Gene ID conversion](#gene-id-conversion-1)

```{r eval=FALSE}
id <-  c("Cyp2c23", "Fhit", "Gal3st2b", "Trp53", "Tp53")
sym2ens <- transId(id ,transTo = "ensembl", org = "mouse")
# keep unique records
sym2ensent <- transId(id ,transTo = c("ensembl","entrez"), org = "mouse", unique = TRUE)
```

## [Protein ID conversion](#protein-id-conversion-1)

```{r eval=FALSE}
id <-  c("P27364", "P14046", "P15650", "P16638", "P24483")
protein2sym <- transId(id ,transTo = "sym", org = "rat")
```

## [Probe ID conversion](#probe-id-conversion-1)


```{r eval=FALSE}
id <-  c('205647_at','204857_at','201394_s_at','210377_at')
probe2sym <- transProbe(id, transTo = 'sym',org = 'hs')
```

## [ORA GO analysis](#ora-go-analysis-1)

> ORA stands for "over-representation analysis".

```{r eval=FALSE}
id <- c("TP53","BRCA1", "PD1", "PDL1")
ego <- genGO(id,
  org = "human",
  ont = "bp",
  pvalueCutoff = 0.01,
  qvalueCutoff = 0.01
)
```

## [ORA KEGG analysis](#ora-kegg-analysis-1)

```{r eval=FALSE}
id <- c("TP53","BRCA1", "PD1", "PDL1","BRCA2","TET2")
keg <- genKEGG(id,
  org = "human",
  pvalueCutoff = 0.05,
  qvalueCutoff = 0.05
)
```

## [FCS GSEA analysis](#functional-class-scoring-1)

> FCS stands for "functional-class-scoring".

`geneList` is an order ranked gene list in decreasing order. Gene inside can be Entrez, Ensembl or symbol.

```{r eval=FALSE}
data(geneList, package = "genekitr")
gse <- genGSEA(
  genelist = geneList,
  org = "human",
  category = "H"
)
```

## [Plot GO and KEGG](#plot-go-kegg-1)

Supported types include: `bar`, `wego`, `bubble`, `dot`, `lollipop`, `geneheat`, `genechord`,
`network`, `gomap`, `goheat`, `gotangram`, `wordcloud`, `upset` and `two-group barplot`

```{r eval=FALSE}
plotEnrich(ego, plot_type = "dot")
plotEnrich(keg, plot_type = "bar")
plotEnrich(ego, plot_type = "geneheat")
plotEnrich(ego, plot_type = "network", scale_ratio = 0.5)
```


## [Plot GSEA](#plot-gsea-1)

Supported types include: `volcano`, `classic`, `fgsea`, `ridge` and `bar`

```{r eval=FALSE}
plotGSEA(gse, plot_type = "volcano", show_pathway = 3)

genes <- c("MET", "TP53", "PMM2")
plotGSEA(gse, plot_type = "classic", show_pathway = pathways, show_gene = genes)

plotGSEA(gse, plot_type = "fgsea", show_pathway = 3)
```

## [Plot Venn](#plot-others-1)

```{r eval=FALSE}
set1 <- paste0(rep("gene", 100), sample(1:1000, 100))
set2 <- paste0(rep("gene", 100), sample(1:1000, 100))
set3 <- paste0(rep("gene", 100), sample(1:1000, 100))

# classic venn plot
plotVenn(list(gset1 = set1, gset2 = set2, gset3 = set3), use_venn = TRUE)
# UpSet plot
plotVenn(list(gset1 = set1, gset2 = set2, gset3 = set3), use_venn = FALSE)
```

## [Plot DEG Volcano](#plot-others-1)

```{r eval=FALSE}
data(deg, package = "genekitr")
plotVolcano(deg, "p.adjust",
  remove_legend = T,
  show_gene = c("CD36", "DUSP6", "IER3", "CDH7")
)
```



