# (PART\*) Part I: Search Information {-}


# Search gene metadata {#search-gene-metadata}

```{r include=FALSE}
library(knitr)
opts_chunk$set(message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE, cache=TRUE)
library(genekitr)
library(dplyr)
```

If users have groups of genes to check their names, location, GC content etc., they can use `genInfo()` function to quickly check.

## Supported organisms {#geninfo-supported-species}

Support 196 vertebrate species, 120 plant species and 2 bacteria species

```{r geninfo-orgname, warning=FALSE}
# install.packages('DT')
library(genekitr)
DT::datatable(ensOrg_name, options = list(pageLength = 10))
```


A user could select an organism name from the `latin_short_name` column. The common name is acceptable for popular research species (e.g., human, mouse, rat, fly, zebrafish, worm, chicken). Take human as an example; its official Latin short name is "happens," but hg," "hsa," "hs," or "human" are also acceptable.

## Basic usage

There are only three arguments in `genInfo()`:

- `id`: gene id (symbol, Entrez or Ensembl) or protein id

- `org`: organism name

- `unique`: `TRUE` or `FALSE`. Commonly, one gene could have many other types of matched records. For example, the human `HBD` gene has three matched Entrez IDs: `r Entrez("3045")`, `r Entrez("85349")`, and `r Entrez("100187828")`. If `TRUE,` only return a one-to-one match result with minimal NAs.

```{r geninfo-basic-research, warning=FALSE}
id <- "HBD"
uniq_res <- genInfo(id, org = 'hs', unique = TRUE)
all_res <- genInfo(id, org = 'hs', unique = FALSE)
uniq_res[1:4]
all_res[1:4]

colnames(all_res)
```

## 




## Useful tricks

User can count organism gene types:

```{r geninfo-biotypes, warning=FALSE}
# you can try other organisms
genInfo(org = 'hs') %>% {table(.$gene_biotype)}
```







