# (PART\*) Part II: ID Conversion {-}

```{r include=FALSE}
library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE, eval = TRUE, echo = TRUE, cache = TRUE)
library(genekitr)
library(clusterProfiler)
library(dplyr)
```

# Gene ID conversion {#gene-id-conversion-1}

If you have worked on genomics analysis before, you may have spent a lot of time converting gene identifiers from one type to another.

> For example, gene set enrichment analysis is very common in the omics downstream process and is mainly based on MSigDB. According to its [release note](https://software.broadinstitute.org/cancer/software/gsea/wiki/index.php/MSigDB_v7.0_Release_Notes), **since MSigDB 7.0, identifiers for genes are mapped to their HGNC approved Gene Symbol and NCBI Gene ID through annotations extracted from Ensembl's BioMart data service**, and are updated at each MSigDB release with the latest available version of Ensembl. Currently, MSigDB 7.5 has updated human gene annotations to Ensembl v105.

Hence, ID conversion helps you stay up to date with the latest annotations.

The most common gene ID types are **symbol** (e.g. TP53), **NCBI Entrez** (e.g. 7157), and **Ensembl** (e.g. ENSG00000141510). ID conversion among these different types is very common, and there are two main concerns: performance and flexibility.


## Current tools {#convertion-current-tools}

Currently, various tools are available, but different methods have their own drawbacks:

- `r Biocpkg("biomaRt")`: This R package is frequently used to access the up-to-date Ensembl database. However, it has to access the online database every time, and there are too many built-in parameters to configure.
- `r Biocpkg("clusterProfiler")`: This R package depends on Bioconductor annotation packages to extract NCBI gene information. NCBI updates its database daily, while Bioconductor packages are updated roughly every six months. Besides, it only supports 20 organisms, and users need to download and load the annotation data before mapping.
- [gProfiler](https://biit.cs.ut.ee/gprofiler/convert): This web server supports 40 types of IDs across more than 60 species. However, it only supports one-to-one conversion — for example, users cannot convert a gene symbol to both Entrez and Ensembl IDs at the same time. Besides, the output result is not very convenient for further processing.
- [DAVID](https://david.ncifcrf.gov/conversion.jsp): The web server interface is not very user-friendly.
- [UniProt](https://www.uniprot.org/uploadlists/): This large protein database supports almost all species, but users need to specify the input ID type, and it only converts to protein IDs.
- [GIDcon](http://resource.ibab.ac.in/GIDCON/geneid/home.html): This web server only covers human, mouse, and rat. Users must separate gene IDs with commas when running a batch query, and converting just five symbols can take up to 8 seconds.


To address the problems above, **genekitr combines both the Ensembl and NCBI databases** to provide more comprehensive results while optimizing running speed.

It also supports **195 vertebrate species, 120 plant species, and two bacterial species** (see also [session 1.1](#geninfo-supported-species)).

## Basic usage {#transid-basic-usage}

`transId` provides five arguments:

- `id`: gene ID (symbol, Entrez, or Ensembl), protein ID (see also [session 4.2](##protein-id-conversion)), or microarray ID (see also [session 5.1](##probe-id-conversion))

- `transTo`: the target type for conversion. You can select **more than one** from "symbol", "entrez", "ensembl", or "uniprot". Abbreviations are also acceptable (e.g., "ens" works for "ensembl", but "en" will raise an error because it cannot be distinguished from "entrez")

- `org`: organism name, default is human

- `unique`: `TRUE` or `FALSE` (default). If `TRUE`, only one matched ID is returned.

- `keepNA`: `TRUE` or `FALSE` (default). Determines whether to keep unmatched IDs (as NA) or remove them.

```{r transid_basic_usage, warning=FALSE}
# human genes
id <- c("AKT3", "SSX6P", "FAKE_ID", "BCC7")
transId(id, transTo = "ens")

# fruit fly genes
transId(
  id = c("10178780", "10178777", "10178786", "10178792"),
  transTo = "sym", org = "fly"
)
```

In the human example above, because the NA was removed, the result order is slightly different from the input order.

If you want to **keep the same order as the input**, you can set `keepNA = TRUE`:
```{r}
transId(id, transTo = "ens", keepNA = TRUE)
```

If you want to **get a one-to-one mapping**, you can set `unique = TRUE`, and the result with the most comprehensive information will be returned:

```{r}
transId(id, transTo = "ens", keepNA = TRUE, unique = TRUE)
```


## Features {#transid-features}

### f1: convert to several types at once {#transid-feature-1}

`transTo` can accept more than one type:

```{r}
transId(id, transTo = c("ens", "ent"))
```

### f2: distinguish between symbol and alias {#transid-feature-2}

One common problem with the [current tools](#convertion-current-tools) mentioned above is confusion between gene symbols and gene aliases.

A gene alias is an alternative name for the official gene symbol — for example, "BCC7" is actually an alias for "TP53". Sometimes a gene alias is even more widely used than the official symbol. For instance, "PDL1" is actually an alias for "CD274".

Thanks to the combination of NCBI and Ensembl metadata, `transId` can easily recognize gene symbols and aliases by setting `transTo = 'symbol'`.

```{r}
transId(c("BCC7", "TP53", "PD1", "PDL1", "TET2"), "sym")
```

### f3: no need to worry about Ensembl versions {#transid-feature-3}

An Ensembl stable ID consists of five parts: ENS(species)(object type)(identifier).(version).
For example, `ENSG00000141510.11` refers to TP53, but the version number `.11` at the end can confuse many tools.
With `transId`, however, you don't need to worry about this issue anymore.

```{r}
transId("ENSG00000141510.11", "symbol")
```


## Exercise 1: DEG result {#transid-ex1-data}
The built-in example data contains differentially expressed genes from [GSE42872](https://www.ncbi.nlm.nih.gov//geo/query/acc.cgi?acc=GSE42872).

```{r}
data(deg, package = "genekitr")
table(deg$change)
id <- deg$entrezid
length(id)
```


### `clusterProfiler::bitr`  {#transid-ex1-bitr}

Users need to specify the input ID type and the target type from `columns(org.Hs.eg.db)`.

```{r}
# You must first install org.Hs.eg.db
# BiocManager::install('org.Hs.eg.db')
library(org.Hs.eg.db)
library(clusterProfiler)
AnnotationDbi::columns(org.Hs.eg.db)
bitr_res <- bitr(id, fromType = "ENTREZID", toType = "SYMBOL", OrgDb = "org.Hs.eg.db")
bitr_sym <- unique(bitr_res$SYMBOL)
length(bitr_sym)
```

### `biomaRt::getBM`  {#transid-ex1-biomart}

The initialization process depends mainly on Ensembl's web access speed, and it usually takes tens of seconds.

```{r eval=FALSE}
# initializing...
ensembl <- biomaRt::useEnsembl(
  biomart = "genes",
  dataset = "hsapiens_gene_ensembl"
)
# conversion
biomart_res <- biomaRt::getBM(
  attributes = c("entrezgene_id", "external_gene_name"),
  filters = c("entrezgene_id"),
  mart = ensembl,
  values = id
)
```


```{r include=FALSE}
load("data/biomart_res.rda")
```

```{r}
biomart_sym <- unique(biomart_res$external_gene_name) %>%
  stringi::stri_remove_empty()
length(biomart_sym)
```


### `genekitr::transId` {#transid-ex1-transid}
```{r}
transid_res <- genekitr::transId(id, transTo = "sym", unique = T)
transid_sym <- unique(transid_res$symbol)
length(transid_sym)
```


### compare the three results {#transid-ex1-compare}

```{r}
plotVenn(list(
  bitr = bitr_sym,
  transid = transid_sym,
  biomart = biomart_sym
))
```

#### **`bitr` vs `transId`**

There are 117 genes found only in the `bitr` result:
```{r}
bitr_sym[!bitr_sym %in% transid_sym] %>% head(5)
```

Let's take a look at the first one, LINC00337, which is converted from gene `r Entrez("148645")`:
```{r}
bitr_res %>%
  filter(SYMBOL %in% "LINC00337") %>%
  pull(ENTREZID)
```


When we search this on NCBI, we find that the official symbol is now "ICMT-DT", while "LINC00337" is just an alias. Looking back at the `transId` result, it stays up to date with the NCBI website.

```{r}
transid_res %>%
  filter(input_id %in% "148645") %>%
  pull(symbol)
```

Besides, there are 118 unique symbols found only in the `transId` result:
```{r}
transid_sym[!transid_sym %in% bitr_sym] %>% head(5)
```

The gene `ICMT-DT` mentioned above is one of them.


#### **`biomart` vs `transId`**

There are 13 genes found only in the `biomart` result:

```{r}
biomart_sym[!biomart_sym %in% transid_sym] %>% head()
biomart_res %>%
  filter(external_gene_name %in% "KLRK1") %>%
  pull(entrezgene_id)
```

When we search NCBI for gene `r Entrez(100528032)`, we find that the official symbol is "KLRC4-KLRK1". Looking back at both the `transId` and `bitr` results, they match the NCBI website.

```{r}
bitr_res %>%
  filter(ENTREZID == "100528032") %>%
  pull(SYMBOL)
transid_res %>%
  filter(input_id %in% "100528032") %>%
  pull(symbol)
```

## Exercise 2: mixture of gene symbols and aliases {#transid-ex2-data}

Compared with Entrez and Ensembl IDs, gene symbols and aliases are much harder to distinguish. In real scientific practice, the impact of confusing the two can be significant.

[Takehara et al. (2018)](https://pubmed.ncbi.nlm.nih.gov/29731861/) had to retract their published paper simply because the authors mistakenly used "TAZ" as the actual research target instead of "WWTR1", where "TAZ" is actually an alias for "WWTR1".

Similar issues have been reported in the paper [The risks of using unapproved gene symbols](https://www.sciencedirect.com/science/article/pii/S0002929721003402). Of course, this problem is not limited to human research alone. So we must stay alert to the risks posed by gene aliases.


### **`HGNChelper` vs `transId`**

`r CRANpkg("HGNChelper")` is mainly designed for human research, so here we'll use a simple human gene list as an example.

```{r}
id <- c("TBET", "B220", "BCC7", "PD1", "PDL1")

transid_res <- genekitr::transId(id, transTo = "sym", unique = FALSE)
transid_res

hgnc_res <- HGNChelper::checkGeneSymbols(id)
hgnc_res
```

The first three genes are omitted in the `HGNChelper` result.

### **`GeneSymbolThesarus` vs `transId`**

`GeneSymbolThesarus` is a function from `r CRANpkg("Seurat")` that looks up current gene symbols from the HUGO Gene Nomenclature Committee (HGNC). Since it searches online, it tends to be relatively slow.

```{r message=TRUE}
library(Seurat)
start <- Sys.time()
seurat_res <- Seurat::GeneSymbolThesarus(symbols = id)
end <- Sys.time()
(end - start)

seurat_res
```

Seurat only returns a single record here.

### **`alias2Symbol` vs `transId`**

`alias2Symbol` is a function from `r Biocpkg("limma")` that relies on Bioconductor's organism annotation packages.

```{r}
limma_res <- limma::alias2Symbol(id, species = "Hs")
# compare with the transId result
id
limma_res
transid_res$symbol
```

It seems that the `alias2Symbol` result does not preserve the original input order, which could easily confuse users. In contrast, `transId` always keeps the original order (and if you want a one-to-one mapping, simply set `unique = TRUE`).
