# Protein ID conversion {#protein-id-conversion-1}

```{r include=FALSE}
library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE, eval = TRUE, echo = TRUE, cache = TRUE)
library(genekitr)
library(dplyr)
```

If you have handled proteomics or metabolomics data (e.g. mass spectrometry), you may need to convert protein IDs to gene IDs before downstream analysis.

## Example data {#protein-example-data}

For example, in this data article [Dataset from proteomic analysis of rat, mouse, and human liver microsomes and S9 fractions](https://www.sciencedirect.com/science/article/pii/S2352340915000189#ec0005), we could get [rat proteomic analysis data](https://ars.els-cdn.com/content/image/1-s2.0-S2352340915000189-mmc1.zip)


Firstly, take a look at the data:

```{r include=FALSE}
load("data/rat_pro.rda")
```


```{r}
colnames(rat_prodata)
head(rat_prodata, 5)
```

The second column is the rat protein ID, and the third one is the gene name.

## ID conversion {#protein-id-conversion-2}

Here we will convert protein IDs to gene symbols and compare them with the original data.

```{r}
protein_id <- rat_prodata[, "Acc"]
gname <- transId(protein_id, "symbol", "rat", unique = T)

new_res <- merge(rat_prodata, gname,
  by.x = "Acc", by.y = "input_id", all.x = T
) %>%
  dplyr::relocate(symbol, .after = Gene)

head(new_res, 5)
```

Maybe you have noticed that some gene names are different from the original data. For example, the second one: `r Uniprot("A0JPQ8")` was previously called `Alkmo`, but its current name is `Agmo`.

Some proteins have been assigned new gene names as shown above, while others currently have no matched name at all.

```{r}
old_name <- unique(new_res$Gene) %>%
  stringr::str_remove_all("_RAT")
new_name <- unique(toupper(new_res$symbol))
new_res[which(is.na(new_res$symbol)), ] %>%
  dplyr::select(Acc, Gene, symbol)
```

We could see the result above: for the protein ID `r Uniprot("P01835")`, the gene name has been replaced with `NA`, indicating no current match was found.

The Venn plot shows that a number of protein IDs have had their corresponding gene names changed. If users continue to use the old gene names, they may miss out on the latest gene information, or even obtain inaccurate enrichment results.

```{r}
plotVenn(list(old_name = old_name, new_name = na.omit(new_name)))
```
