# Probe ID conversion {#probe-id-conversion}

```{r include=FALSE}
library(knitr)
opts_chunk$set(message=FALSE, warning=FALSE, eval=TRUE, echo=TRUE, cache=TRUE)
library(genekitr)
library(dplyr)
```

Microarray probes are used to measure the expression level of many genes. After getting microarray data, the first important thing before downstream analysis is probe id conversion.

`genekitr::transProbe` is built on [Ensembl up-to-date annotation datasets](https://www.ensembl.org/info/genome/microarray_probe_set_mapping.html) and could join together with other genekitr functions elegantly.

Supported platforms can be listed by:

```{r}
DT::datatable(hsapiens_probe_platform, options = list(pageLength = 10))
```

## Probe id conversion {#probe-id-conversion}

The user only needs to input probe id, organism, and wanted conversion type. Leave the rest to genekitr! It will keep the same order with input probes.

```{r}
data(deg, package = "genekitr")
probes <- deg$probe_id

res <- transProbe(probes,"symbol",org = "human")
head(probes); head(res)
```


## Convert to many types {#convert-probe-types}

User could define `transTo` as many types:

```{r}
res2 <- transProbe(probes, c("symbol","ensembl"), org = "human")
head(res2)
```





