# ORA human disease analysis {#ora-human-disease-analysis-1}

```{r include=FALSE}
library(knitr)
opts_chunk$set(message = FALSE, warning = FALSE, eval = TRUE, echo = TRUE, cache = TRUE)
library(genekitr)
```

## Available arguments {#ora-aavailable-arguments}

- `id`: gene id of Entrez, Ensembl, symbol or Uniprot
- `type`: choose one from `do` (Disease Ontology), `ncg` (Network of Cancer Gene) and `dgn` (DisGeNET)
- `group_list`: gene set comparison groups, default is NULL.
- `pAdjustMethod`: choose one from "holm", "hochberg", "hommel", "bonferroni", "BH", "BY", "fdr", "none".
- `pvalueCutoff`: pvalue cutoff, default is 0.05.
- `qvalueCutoff`: adjusted pvalue cutoff, default is 0.05.
- `minGSSize`: minimal size of each geneSet for analyzing, default is 10.
- `maxGSSize`: maximal size of each geneSet for analyzing, default is 500. Program will choose gene set in range [minGSSize,maxGSSize] to analyze
- `universe`: Background genes of Entrez, Ensembl, symbol or Uniprot. Default is NULL. 


## Disease Ontology (DO) {#ora-do}

The [Disease Ontology](https://disease-ontology.org/) has been developed as a standardized ontology for human disease with the purpose of providing the biomedical community with consistent, reusable and sustainable descriptions of human disease terms.

```{r}
data(geneList, package = "genekitr")
id <- names(geneList)[abs(geneList) > 1]
edo <- genORA(id, type = "do")
head(edo)

# if we input ensembl genes
ens_id <- transId(id,'ens',unique = TRUE)[,2]
ens_edo <- genORA(ens_id, type = "do")
head(ens_edo)
```

## Network of Cancer Gene (NCG) {#ora-ncg}

[NCG](http://ncg.kcl.ac.uk/) is a manually curated collection of cancer genes, healthy drivers and their properties.

```{r}
encg <- genORA(names(geneList), type = "ncg",
               pvalueCutoff = 1,qvalueCutoff = 1)
head(encg)
```

## DisGeNET (DGN) {#ora-dgn}

[DisGeNET](https://www.disgenet.org/) is a discovery platform containing one of the largest publicly available collections of genes and variants associated to human diseases.

DisGeNET integrates data from expert curated repositories, GWAS catalogues, animal models and the scientific literature. DisGeNET data are homogeneously annotated with controlled vocabularies and community-driven ontologies. Additionally, several original metrics are provided to assist the prioritization of genotype–phenotype relationships.

```{r}
edgn <- genORA(id, type = "dgn", 
               pvalueCutoff = 0.01, 
               qvalueCutoff = 0.01)
```


## Groups comparison

All these data sets support gene group comparison, just pass the `group_list` to function.

Take disease ontology as example:

```{r}
id <- c(head(names(geneList), 1000), tail(names(geneList), 1000))
group <- list(
  group1 = c(rep("up", 1000), rep("down", 1000)),
  group2 = c(rep("A", 1300), rep("B", 700))
)
gedo <- genORA(id, type = "do", group_list = group)
head(gedo)
```


