# Plot others {#plot-others-1}

```{r include=FALSE}
library(knitr)
library(ggplot2)
library(igraph)
library(ggraph)
library(patchwork)
library(genekitr)
```

In this part, we will introduce `plotVenn` and `plotVolcano`.

## Venn Plot

Venn plot needs a list of gene sets, for example:

```{r}
set1 <- paste0(rep("gene", 50), sample(1:1000, 50))
set2 <- paste0(rep("gene", 60), sample(1:1000, 60))
set3 <- paste0(rep("gene", 70), sample(1:1000, 70))
set4 <- paste0(rep("gene", 80), sample(1:1000, 80))
set5 <- paste0(rep("gene", 100), sample(1:1000, 100))

# three groups
sm_gene_list <- list(gset1 = set1, gset2 = set2, gset3 = set3)

# five groups
la_gene_list <- list(
  gset1 = set1, gset2 = set2, gset3 = set3,
  gset4 = set4, gset5 = set5
)
```

`plotVenn` supports two types: Venn (default) and UpSet. Besides, it also supports modification of text size, alpha degree, colors and other parameters from `plot_theme`.

(ref:vennScap) Venn plot

(ref:vennCap) **Venn plot** 

```{r venn, fig.height=5, fig.width=6, fig.cap="(ref:vennCap)", fig.scap="(ref:vennScap)", dpi=300}
plotVenn(sm_gene_list,
         use_venn = TRUE,
         color = ggsci::pal_lancet()(3),
         alpha_degree = 1,
         main_text_size = 1.5,
         border_thick = 0
)
```

UpSet plot is better for over four groups of genes:

(ref:upsetVennScap) UpSet plot

(ref:upsetVennCap) **UpSet plot** 

```{r upsetVenn, fig.height=4, fig.width=8, fig.cap="(ref:upsetVennCap)", fig.scap="(ref:upsetVennScap)", dpi=300}
plotVenn(la_gene_list,
         use_venn = FALSE,
         main_text_size = 15,
         legend_text_size = 8,
         legend_position = 'left'
)
```

## Volcano Plot

After finishing differential expression analysis, we could get differentially expressed genes (DEG).
Volcano plot is widely used for these up/down regulated genes visualization.

The example data is micro array differential expression analysis:

```{r}
data(deg, package = "genekitr")
head(deg,5)
```


`plotVolcano` mainly uses a DEG result which should include gene id, fold change and statistic value. 

It also has other parameters:

- `stat_metric`: choose from "pvalue" and "p.adjust"
- `stat_cutoff`: cutoff for significant statistic value, default is 0.05
- `logFC_cutoff`: cutoff for log2 (fold change), default is 1 which is actually two fold changes
- `up_color`: color for up-regulated genes
- `down_color`: color for down-regulated genes
- `show_gene`: show some gene names, default is none. Suggest renaming the column with these genes as "gene".
- `dot_size`: volcano dot size

(ref:degVolcanoScap) Volcano plot

(ref:degVolcanoCap) **Volcano plot** 

```{r degVolcano, fig.height=5, fig.width=6, fig.cap="(ref:degVolcanoCap)", fig.scap="(ref:degVolcanoScap)", dpi=300}
plotVolcano(deg, "p.adjust", remove_legend = T, dot_size = 3)
```

If we want to show some genes:

(ref:degVolcano2Scap) Volcano plot with genes

(ref:degVolcano2Cap) **Volcano plot with genes.** 

```{r degVolcano2, fig.height=5, fig.width=6, fig.cap="(ref:degVolcano2Cap)", fig.scap="(ref:degVolcano2Scap)", dpi=300}
plotVolcano(deg, 
            stat_metric = "p.adjust",
            remove_legend = T,
            show_gene = c("CD36", "DUSP6", "IER3", "CDH7")
)
```

> Trick for increasing text size:

User could hide all x/y-axis labels and raise `main_text_size` to enlarge gene text, then export the figure into PPT to add labels.

```{r}
plotVolcano(deg, 
            "p.adjust",
            remove_legend = T,
            remove_main_text = T,
            show_gene = c("CD36", "DUSP6", "IER3", "CDH7"),
            main_text_size = 15, 
            dot_size = 2.5
)
```




